Better Splitting Algorithms for Parallel Corpus Processing

### 2012-08-02 doc/latex/paper.tex
Split figures into subfigures, per request by Reviewer 1.
Also improves drawing of bar graphs, per request by Reviewer 1.

\begin{figure*}[p]
\centering
%\subfloat[Decoding times in \textbf{seconds} for standard decoder configured using a 5-gram language model.]{
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={160,150,140,130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Seconds),
ymin=0,
%legend style={
%	at={(0.5,-0.15)},
%	anchor=north,legend columns=-1
%},
ybar=0pt,
%bar width=3pt,
bar shift=0pt
]
%
\input{ngram/ngram.naive-Maximum_times}
\input{ngram/ngram.histogram-Maximum_times}
\input{ngram/ngram.words-Maximum_times}
\input{ngram/ngram.time-Maximum_times}
%
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
%}
\caption{\fontspec{TeX Gyre Pagella}Decoding times (in {\textbf{seconds}}) for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}) for a phrase-based decoder configured using a 5-gram language model.}
\label{fig:maxspeed}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%
%\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[p]
\centering
%\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.]{
\begin{tikzpicture}%[scale=2]
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Hours),
ymin=0,
ybar=0pt,
bar shift=0pt
]
%
\input{hhmm/hhmm.naive-Maximum_times}
\input{hhmm/hhmm.histogram-Maximum_times}
\input{hhmm/hhmm.words-Maximum_times}
\input{hhmm/hhmm.time-Maximum_times}
%
%\legend{Naive,Histogram,Words,Oracle}
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
%}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\caption{\fontspec{TeX Gyre Pagella}Decoding times (in \textbf{hours}) for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}) for a phrase-based decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.}
\label{fig:maxspeedslm}
\end{figure*}

\begin{figure*}
\centering
%\subfloat[Decoding times in \textbf{seconds} for standard decoder configured using a 5-gram language model.]{
\input{ngram/ngram.max-min-table.tex}
%} \\
\caption{\fontspec{TeX Gyre Pagella}Decoding times in \textbf{seconds} for the fastest (min) and slowest (max) decoding jobs (for a phrase-based decoder configured using a 5-gram language model) when a translation task is split into $n$ decoding jobs. \textit{Italics} indicate balanced task times. \textbf{Bold} indicates fastest max time at that split.}
\label{fig:minmaxspeed}
\end{figure*}

\begin{figure*}
%\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) addition to a 5-gram language model.]{
\input{hhmm/hhmm.max-min-table.tex}
%}
\caption{\fontspec{TeX Gyre Pagella}Decoding times in \textbf{hours} for the fastest (min) and slowest (max) decoding jobs (for a phrase-based decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model) when a translation task is split into $n$ decoding jobs. \textit{Italics} indicate balanced task times. \textbf{Bold} indicates fastest max time at that split.}
\label{fig:minmaxspeedhhmm}
\end{figure*}

@@

To observe the effects of splitting algorithms on decoding speed, we translated Urdu-English data using Moses in a parallel computing cluster, distributing work using the Sun Grid Engine. We ran two decoding setups: a standard configuration using a 5-gram language model, and a much slower syntactic LM configuration following \citet{schwartzetal11}. Using Algorithm \ref{alg:naive}, the runtimes of the slowest of $n$ translation jobs in each configuration is illustrated in Figures \ref{fig:maxspeed} and \ref{fig:maxspeedslm} for various values of $n$. Figures \ref{fig:minmaxspeed} and \ref{fig:minmaxspeedhhmm} show that in all cases, there is a significant difference between the fastest and slowest translation job, leading to computational slack time (Figure \ref{fig:slack}).

@@

In examining these results, we observe that the slack time results primarily from situations where some jobs are assigned a disproportionate number of short sentences; because short sentences take less time to translate than long sentences, such jobs finish much faster than jobs that are assigned many longer sentences. To remedy this imbalance, we propose Algorithm \ref{alg:histogram}. This technique examines the word lengths of each sentence prior to splitting the data into jobs. Sentences are sorted according to length, then assigned in turns to jobs. This results in the sentence length histograms for each job being approximately equal. 
%
%
%
%In nearly all translation models, each sentence is translated independently 
%
%\todo[inline]{Describe Algorithm \ref{alg:words}. \\ \ \\ \ \\ \ \\ \ \\ \ \\ \ }
%
%
In Figure \ref{fig:minmaxspeed}, we observe that the maximum runtimes using Algorithm \ref{alg:histogram} fail to improve over the runtimes using Algorithm \ref{alg:naive} for a standard Moses configuration for small values of $n$. But for values of $n>8$, and for all values of $n$ using the slow syntactic language model, Figure \ref{fig:minmaxspeedhhmm} shows that Algorithm \ref{alg:histogram} results in significant decreases in total decoding time over the naive Algorithm \ref{alg:naive}.

@@

\begin{figure*}
\centering
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Slack \textbf{CPU-Seconds} for standard decoder configured using a 5-gram language model.]{
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={2,4,8,16,32,64,128,256,512},
scaled y ticks=false,
ytick={2000,1800,1600,1400,1200,1000,800,600,400,200,100},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Slack CPU-Seconds,
ymin=0,
legend pos=north west,
ybar=0pt,
%bar width=3pt,
bar shift=0pt
]
%
\input{ngram/ngram.naive-slack_times}
\input{ngram/ngram.histogram-slack_times}
\input{ngram/ngram.words-slack_times}
\input{ngram/ngram.time-slack_times}
%
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%
\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Slack \textbf{CPU-Hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.]{
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={2,4,8,16,32,64,128,256,512},
ytick={3000,2500,2000,1500,1000,750,500,250,100},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Slack CPU-Hours,
ymin=0,
legend pos=north west,
ybar=0pt,
%bar width=3pt,
bar shift=0pt
]
%
\input{hhmm/hhmm.naive-slack_times}
\input{hhmm/hhmm.histogram-slack_times}
\input{hhmm/hhmm.words-slack_times}
\input{hhmm/hhmm.time-slack_times}
%
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\caption{Cumulative slack CPU time for $n$ processing cores when processing a parallel translation task split into $n$ jobs using various splitting algorithms. Slack CPU time is caused when some jobs finish before others. Zero slack time indicates conditions where all jobs complete simultaneously.}
\label{fig:slack}
\end{figure*}

@@

While the most effective algorithm (Algorithm \ref{alg:times}) requires per-sentence decode times from previous decodes, in most realistic settings, Algorithm \ref{alg:words} provides nearly all the benefits of Algorithm \ref{alg:times} in terms of decreased computational slack time while requiring little changes to existing decoding scripts which use the naive Algorithm \ref{alg:naive}.

>>>>

\begin{figure*}[p]
\centering
\subfloat[Decoding times in \textbf{seconds} for decoder configured using a 5-gram language model.]{
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={160,150,140,130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Seconds),
ymin=0,
%enlarge x limits=0.5,
%scale only axis=false,
%axis equal=false,
%legend style={
%	at={(0.5,-0.15)},
%	anchor=north,legend columns=-1
%},
%ybar=0pt,
width=13cm,
height=7cm,
ybar,
bar width=7pt,
cycle list={fill=gray!15!blue!15,fill=gray!50!orange!50,fill=gray!70!green!70,fill=gray!100}
%bar shift=3pt
%ybar interval=0.7
]
%
\input{ngram/ngram.naive-Maximum_times}
\input{ngram/ngram.histogram-Maximum_times}
\input{ngram/ngram.words-Maximum_times}
\input{ngram/ngram.time-Maximum_times}
%
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%\caption{\fontspec{TeX Gyre Pagella}Decoding times (in {\textbf{seconds}}) for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}) for a phrase-based decoder configured using a 5-gram language model.}
%\label{fig:maxspeed}
%\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%

%\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure*}[p]
%\centering
\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.]{
\begin{tikzpicture}%[scale=2]
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Hours),
ymin=0,
%ybar=0pt,
ybar,
bar width=7pt,
%bar shift=0pt,
width=13cm,
height=7cm,
%cycle list name=black white
%colormap name=bluered
%cycle list={fill={blue!80!red!30}, fill=red!66, fill=yellow!33, black}
%cycle list={fill=blue!30!white,fill=red!60!white,fill=brown!90!white,fill=black}
%cycle list={fill=blue!30!white,fill=red!30!white,fill=brown!30!white,fill=grey}
%cycle list={fill=gray!25,fill=gray!50,fill=gray!75,fill=gray!100}
cycle list={fill=gray!15!blue!15,fill=gray!50!orange!50,fill=gray!70!green!70,fill=gray!100}
%cycle list name=color
]
%
\input{hhmm/hhmm.naive-Maximum_times}
\input{hhmm/hhmm.histogram-Maximum_times}
\input{hhmm/hhmm.words-Maximum_times}
\input{hhmm/hhmm.time-Maximum_times}
%
%\legend{Naive,Histogram,Words,Oracle}
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%\caption{\fontspec{TeX Gyre Pagella}Decoding times (in \textbf{hours}) for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}) for a phrase-based decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.}
\caption{\fontspec{TeX Gyre Pagella}Decoding times for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}).}
\label{fig:maxspeed}
\end{figure*}

\begin{figure*}
\centering
\subfloat[Decoding times in \textbf{seconds} for decoder configured using a 5-gram language model.]{
\input{ngram/ngram.max-min-table.tex}
} \\
%\caption{\fontspec{TeX Gyre Pagella}Decoding times in \textbf{seconds} for the fastest (min) and slowest (max) decoding jobs (for a phrase-based decoder configured using a 5-gram language model) when a translation task is split into $n$ decoding jobs. \textit{Italics} indicate balanced task times. \textbf{Bold} indicates fastest max time at that split.}
%\label{fig:minmaxspeed}
%\end{figure*}
%
%\begin{figure*}
\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) addition to a 5-gram language model.]{
\input{hhmm/hhmm.max-min-table.tex}
}
%\caption{\fontspec{TeX Gyre Pagella}Decoding times in \textbf{hours} for the fastest (min) and slowest (max) decoding jobs (for a phrase-based decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model) when a translation task is split into $n$ decoding jobs. \textit{Italics} indicate balanced task times. \textbf{Bold} indicates fastest max time at that split.}
\caption{\fontspec{TeX Gyre Pagella}Decoding times for the fastest (min) and slowest (max) decoding jobs when a translation task is split into $n$ decoding jobs. \textbf{Bold} indicates fastest max time at that split. \textit{Italics} indicate balanced task times, corresponding to zero slack time (see Figure \ref{fig:slack}).}
%\label{fig:minmaxspeedhhmm}
\label{fig:minmaxspeed}
\end{figure*}

@@

To observe the effects of splitting algorithms on decoding speed, we translated Urdu-English data using Moses in a parallel computing cluster, distributing work using the Sun Grid Engine. We ran two decoding setups: a standard configuration using a 5-gram language model, and a much slower syntactic LM configuration following \citet{schwartzetal11}. Using Algorithm \ref{alg:naive}, the runtimes of the slowest of $n$ translation jobs in each configuration is illustrated in Figure \ref{fig:maxspeed} for various values of $n$. Figure \ref{fig:minmaxspeed} shows that in all cases, there is a significant difference between the fastest and slowest translation job, leading to computational slack time (Figure \ref{fig:slack}).

@@

In examining these results, we observe that the slack time results primarily from situations where some jobs are assigned a disproportionate number of short sentences; because short sentences take less time to translate than long sentences, such jobs finish much faster than jobs that are assigned many longer sentences. To remedy this imbalance, we propose Algorithm \ref{alg:histogram}. This technique examines the word lengths of each sentence prior to splitting the data into jobs. Sentences are sorted according to length, then assigned in turns to jobs. This results in the sentence length histograms for each job being approximately equal. 
%
%
%
%In nearly all translation models, each sentence is translated independently 
%
%\todo[inline]{Describe Algorithm \ref{alg:words}. \\ \ \\ \ \\ \ \\ \ \\ \ \\ \ }
%
%
In Figure \ref{fig:minmaxspeed}a, we observe that the maximum runtimes using Algorithm \ref{alg:histogram} fail to improve over the runtimes using Algorithm \ref{alg:naive} for a standard Moses configuration for small values of $n$. But for values of $n>8$, and for all values of $n$ using the slow syntactic language model, Figure \ref{fig:minmaxspeed}b shows that Algorithm \ref{alg:histogram} results in significant decreases in total decoding time over the naive Algorithm \ref{alg:naive}.

@@

\begin{figure*}
\centering
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Slack \textbf{CPU-Seconds} for decoder configured using a 5-gram language model.]{
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={2,4,8,16,32,64,128,256,512},
scaled y ticks=false,
ytick={2000,1800,1600,1400,1200,1000,800,600,400,200,100},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Slack CPU-Seconds,
ymin=0,
legend pos=north west,
%ybar=0pt,
%bar width=3pt,
bar shift=0pt,
width=13cm,
height=7cm,
ybar,
bar width=5pt,
cycle list={fill=gray!15!blue!15,fill=gray!50!orange!50,fill=gray!70!green!70,fill=gray!100}
]
%
\input{ngram/ngram.naive-slack_times}
\input{ngram/ngram.histogram-slack_times}
\input{ngram/ngram.words-slack_times}
\input{ngram/ngram.time-slack_times}
%
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%
\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Slack \textbf{CPU-Hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.]{
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={2,4,8,16,32,64,128,256,512},
ytick={3000,2500,2000,1500,1000,750,500,250,100},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Slack CPU-Hours,
ymin=0,
legend pos=north west,
%ybar=0pt,
%bar width=3pt,
bar shift=0pt,
width=13cm,
height=7cm,
ybar,
bar width=5pt,
cycle list={fill=gray!15!blue!15,fill=gray!50!orange!50,fill=gray!70!green!70,fill=gray!100}
]
%
\input{hhmm/hhmm.naive-slack_times}
\input{hhmm/hhmm.histogram-slack_times}
\input{hhmm/hhmm.words-slack_times}
\input{hhmm/hhmm.time-slack_times}
%
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\caption{Cumulative slack CPU time for $n$ processing cores when processing a parallel translation task split into $n$ jobs using various splitting algorithms. Slack CPU time is caused when some jobs finish before others. Zero slack time indicates conditions where all jobs complete simultaneously.}
\label{fig:slack}
\end{figure*}

@@

While the most effective algorithm (Algorithm \ref{alg:times}) requires per-sentence decode times from previous decodes, in most realistic settings, Algorithm \ref{alg:words} provides much of the benefit of Algorithm \ref{alg:times} in terms of decreased computational slack time while requiring little changes to existing decoding scripts which use the naive Algorithm \ref{alg:naive}.
